---
title: "Preprocessing access layers"
output: html_document
---

## About

In app v0.3.3 (and prior versions), whether a 100x100m raster is accessible for a recorder to go an visit was an on-the -fly operation done by loading in a variety of shapefiles then doing simple feature operations to determine whether the grid square was within 200m of an accessible feature. It was unfortunantly a bit slow.

See https://github.com/BiologicalRecordsCentre/DECIDE-app/issues/141 for discussion and progress

This script is a proof of concept that we can creates a 100x100 'access score' raster layer that is loaded into the app to speed up nudge generation. This raster is a result of 'pre-processing' the access information we hold.

It does it by working through each of 10km grid squares, loads in data from open street map overpass API, and static files.

In then does uses sf functions to check, primarily `st_is_within_distance` to check whether each grid square (or inactually, the centroid of each grid square) is within some distance of an accessible feature.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
#load packages
library(osmdata) #for using overpass API
library(sf)
library(leaflet) # for making maps for sanity checking
library(raster)
library(rgdal)
library(dplyr)
library(htmlwidgets)

```

Load UK 10k grids and 100x100 raster

```{r}
uk_grid <- st_read('/data/data/DECIDE_constraintlayers/raw_data/UK_grids/uk_grid_10km.shp')
st_crs(uk_grid) <- 27700

# sanity check that the grid squares look like the UK
#plot(st_geometry(uk_grid), add = F, col = 'orange')

# the big raster
raster100 <- raster::stack('/data/data/DECIDE_constraintlayers/environmental_data/100mRastOneLayer.grd')
#plot(raster100)

#get the CRS for when we project back to raster from data ramew
raster_crs <- crs(raster100)

raster100_df <- as.data.frame(raster100, xy=T, centroids=TRUE)[,1:2]

```

Map layers

Offline files

Currently in use for pre-processing:
 - CRow Act 2000
 - OS greenspace
 - RSPB reserve boundaries
 - SSSI

 - Greater London
 - National trust
 - rowmaps_footpath_bridleway
 
Downloaded but not in use:
 - OS roadnetwork
 - Scotland
 
OSM layers



### Function for getting offline files for a 10km grid number

```{r}

#base location
base_location <- '/data/data/DECIDE_constraintlayers/raw_data/'

#folders
file_locations <- c(
  'CRoW_Act_2000_-_Access_Layer_(England)-shp/gridded_data_10km/',
  'OS_greenspaces/OS Open Greenspace (ESRI Shape File) GB/data/gridded_greenspace_data_10km/',
  #'SSSIs/gridded_data_10km/',
  'greater-london-latest-free/london_gridded_data_10km/',
  'rowmaps_footpathbridleway/rowmaps_footpathbridleway/gridded_data_10km/',
  #'RSPB_Reserve_Boundaries/gridded_data_10km/',
  'national_trust/gridded_data_10km/'
)

# function to get the data
get_offline_gridded_data <- function(grid_number){
  returned_files <- list()
  
  for (ii in 1:length(file_locations)){
    file_location <- file_locations[ii]
    all_grids <- list.files(paste0(base_location,file_location))
    
    right_file <- all_grids[grep(paste0("_",grid_number,".shp"), all_grids)]
    
    if (length(right_file)>0){
      #build file path
      file_path <- paste0(base_location,file_location,right_file)
      
      #load in file
      file_shp <- st_read(file_path, quiet = TRUE) %>% st_transform(4326)
      
      #add it to the object that this function returns
      returned_files[[length(returned_files)+1]] <- file_shp
    }
  }
  #return the files
  returned_files
}


#test <- get_offline_gridded_data(1500)



```


## Pre-processing

This loops through 

```{r}
#setting to F resolved some issues but 
sf::sf_use_s2(T)

test_maps <- c(119,320,400,984,1082,1313,runif(100,min = 0, max = 2000) %>% round())
test_maps <- c(runif(100,min = 0, max = 2000) %>% round())

#loop through all of the UK 10k grid squares
for (i in test_maps){
  
  print(paste0("grid number: ",i))
  
  start.time <- Sys.time()
  
  ### Getting OSM data
  
  #make the bounding box for the OSM query
  this_10k_grid <- uk_grid[i,]$geometry #get the 10kgrid
  this_10k_gridWGS84 <- st_transform(this_10k_grid, 4326) # convert to WGS84
  
  #where are we?
  if(T){
    m <- leaflet() %>%
    addTiles() %>%
    addPolygons(data=this_10k_gridWGS84,opacity=1,fillOpacity = 0,weight=2,color = "black")
    m
  }
  
  
  # build the OSM query
  q1 <- opq(bbox = this_10k_gridWGS84) %>%
    add_osm_feature(key = 'highway', value = 'footway') %>% #footways
    add_osm_feature(key = 'designation', value = 'public_footpath') %>% #properly designated public footpaths, england and wales only
    #add_osm_feature(key = 'highway', value = 'bridleway') %>% # bridleways
    add_osm_feature(key = 'designation', value = 'public_bridleway') %>% #properly designated public footpaths, england and wales only
    #add_osm_feature(key = 'highway', value = 'steps') %>% # steps (basically footpaths), probably not worth querying because usually such small features
    add_osm_feature(key = 'highway', value = 'path') %>% # generic paths
    #add_osm_feature(key = 'highway', value = 'track') %>% # gravel tracks etc Excluded becase often private tracks are included
    add_osm_feature(key = 'designation', value = 'byway_open_to_all_traffic') %>% 
    add_osm_feature(key = 'designation', value = 'restricted_byway') %>% 
    add_osm_feature(key = 'designation', value = 'access_land') %>% 
    add_osm_feature(key = 'highway', value = 'residential') %>% # roads with housing so presumably will be accessible
    add_osm_feature(key = 'highway', value = 'unclassified') %>% #really small roads roads (not actually 'unclassified')
    add_osm_feature(key = 'highway', value = 'tertiary') %>% # roads smaller than b roads etc.
    add_osm_feature(key = 'footway', value = 'sidewalk') %>% # pavements if classified separately to roads
    add_osm_feature(key = 'sidewalk', value = 'both') %>% #roads with pavements 
    add_osm_feature(key = 'sidewalk', value = 'left') %>% #roads with pavements
    add_osm_feature(key = 'sidewalk', value = 'right') #roads with pavements
  
  #bad features
  q2 <- opq(bbox = this_10k_gridWGS84) %>%
    add_osm_feature(key = 'landuse', value = 'quarry') %>%
    add_osm_feature(key = 'aeroway', value = 'aerodrome') %>%
    add_osm_feature(key = 'railway', value = 'rail') %>%
    add_osm_feature(key = 'highway', value = 'motorway') %>%
    add_osm_feature(key = 'highway', value = 'trunk') %>%
    add_osm_feature(key = 'landuse', value = 'landfill')
  
  #warning features
  q3 <- opq(bbox = this_10k_gridWGS84) %>%
    add_osm_feature(key = 'landuse', value = 'military')
  
  #water
  q4 <- opq(bbox = this_10k_gridWGS84) %>% 
    add_osm_feature(key = 'natural', value = 'water') %>%
    add_osm_feature(key = 'waterway', value = 'river')
  
  print("loading OSM data")
  #get OSM data
  q1$prefix <- q2$prefix <- q3$prefix <- q4$prefix <- "[out:xml][timeout:10000];\n(\n"
  good_features_data <- osmdata_sf(q1)
  bad_features_data <- osmdata_sf(q2)
  warning_features_data <- osmdata_sf(q3)
  wet_features_data <- osmdata_sf(q4)
  
  ### Getting PROW, Access land etc. data
  
  print("loading offline data")
  offline_good_features_data <- get_offline_gridded_data(i)
  

  # get the bounding box of the 10km grid square
  grid_bb <- st_bbox(this_10k_grid)
  
  #only select the raster squares in the 10km grid square
  raster_this_grid <- raster100_df %>% filter(x > grid_bb$xmin,
                                              x < grid_bb$xmax,
                                              y > grid_bb$ymin,
                                              y < grid_bb$ymax)
  
  #get the easting and northing projection
  projcrs <- crs(this_10k_grid)
  
  # make the raster df into a sd object using the projection but transform it to WGS84 for these operations
  raster_as_sf <- st_as_sf(raster_this_grid,coords = c("x", "y"),crs = projcrs) %>% st_transform(4326)
  
  distance_check_bad <- distance_check_good <- distance_check_warning <- distance_check_water <- rep(0,nrow(raster_this_grid))
  
  print("Checking distances of OSM data")
  
  #see if the grid squares are within each of the good features
  if (!is.null(good_features_data$osm_lines)) {
    distance_check_good <- distance_check_good + rowSums(st_is_within_distance(raster_as_sf,good_features_data$osm_lines,dist = 100,sparse = FALSE))
  }
  
  
  print("Checking distances of offline data")
  # do the same for all the good offline features
  if (length(offline_good_features_data)>0){
    for (ii in 1:length(offline_good_features_data)){
      print(ii)
      
      #sometimes this would error and a fix was to turn off sphyical gemoetry so this trycatch turns it off if there's an error
      tryCatch(
        {
        distance_check_good <- distance_check_good + rowSums(st_is_within_distance(raster_as_sf,offline_good_features_data[[ii]]$geometry,dist = 100,sparse = FALSE))
        
        }, error = function(cond){
        sf::sf_use_s2(F)
          distance_check_good <- distance_check_good + rowSums(st_is_within_distance(raster_as_sf,offline_good_features_data[[ii]]$geometry,dist = 100,sparse = FALSE))
          sf::sf_use_s2(T)
        }
      )
      
    }
  }
  
  print("Checking distances of bad OSM data")
  # do the same for the OSM bad, warning and water features
  #polygons
  if (!is.null(bad_features_data$osm_polygons)){
    distance_check_bad <- distance_check_bad + rowSums(st_is_within_distance(raster_as_sf,bad_features_data$osm_polygons,dist = 25,sparse = FALSE))
  }
  
  # lines
  if (!is.null(bad_features_data$osm_lines)){
    distance_check_bad <- distance_check_bad+  rowSums(st_is_within_distance(raster_as_sf,bad_features_data$osm_lines,dist = 50,sparse = FALSE))
  }
  
  # warning features
  if (!is.null(warning_features_data$osm_polygons)){
    distance_check_warning <- distance_check_warning + rowSums(st_is_within_distance(raster_as_sf,warning_features_data$osm_polygons,dist = 25,sparse = FALSE))
  }
  
  #water
  if (!is.null(wet_features_data$osm_polygons)){
    distance_check_water <- rowSums(st_is_within_distance(raster_as_sf,wet_features_data$osm_polygons,dist = 0,sparse = FALSE))
    buffered_water_points <- raster_as_sf[distance_check_water>0,] %>% st_buffer(dist=50)
    
    
    points_in_water <- st_covered_by(buffered_water_points,wet_features_data$osm_polygons,sparse = F) %>% rowSums()
    rownames(buffered_water_points)[points_in_water>0]
    
    distance_check_water <- 0
    
    distance_check_water[rownames(raster_as_sf) %in% rownames(buffered_water_points)[points_in_water>0]] <- 1
  }
  
  raster_as_sf$access <- distance_check_good
  raster_as_sf$no_go <- distance_check_bad
  raster_as_sf$warning <- distance_check_warning
  raster_as_sf$water <- distance_check_water
  
  raster_as_sf$composite <- 0.5 # neutral
  raster_as_sf$composite[raster_as_sf$access>0] <- 1 #go
  raster_as_sf$composite[raster_as_sf$warning>0] <- 0.75 # warning
  raster_as_sf$composite[raster_as_sf$no_go>0] <- 0 #no go
  raster_as_sf$composite[raster_as_sf$water>0] <- 0.25 #water

  
  print("Making map")
  # View this  on a leaflet map
  m <- leaflet() %>%
    addTiles() %>%
    #addPolylines(data=good_features_data$osm_line,opacity=1,weight=2) %>%
    addPolygons(data=this_10k_gridWGS84,opacity=1,fillOpacity = 0,weight=2,color = "black") %>%
    addCircles(data = raster_as_sf %>% filter(composite==1),radius = 50,weight=0,color = "blue") %>%
    addCircles(data = raster_as_sf %>% filter(composite==0),radius = 50,weight=0,color = "red") %>%
    addCircles(data = raster_as_sf %>% filter(composite==0.75),radius = 50,weight=0,color = "orange") %>%
    addCircles(data = raster_as_sf %>% filter(composite==0.25),radius = 50,weight=0,color = "black",fillOpacity = 0.8) %>%
    addPolygons(data=wet_features_data$osm_polygons,opacity=1,fillOpacity = 0.3,weight=2,color = "black")
  
  saveWidget(m, file=paste0("maps/access_map_grid_",i,".html"))
  
  print("Saving data")
  # add the reaster info
  raster100_df[raster100_df$x > grid_bb$xmin & raster100_df$x < grid_bb$xmax & raster100_df$y > grid_bb$ymin & raster100_df$y < grid_bb$ymax,"access"] <- raster_as_sf$composite
  
  raster100_to_save <- raster100_df[raster100_df$x > grid_bb$xmin & raster100_df$x < grid_bb$xmax & raster100_df$y > grid_bb$ymin & raster100_df$y < grid_bb$ymax,]
  saveRDS(raster100_to_save,file = paste0("/data/data/DECIDE_constraintlayers/processed_data/access_raster_grid",i,".RDS"))
  
  #save progress periodically
  # if ((i %% 25) == 0){
  #   saveRDS(raster100_df,file = "/data/data/DECIDE_constraintlayers/processed_data/access_raster.RDS")
  # }
  
  end.time <- Sys.time()
  time.taken <- end.time - start.time
  print(paste0("Time taken: ",time.taken))
}

if(F){
  # rebuild the raster
  raster100_with_access_info <- rasterFromXYZ(raster100_df,crs = raster_crs)
  raster100_with_access_info$access
}


```

 