---
title: "Preproccessing access layers v2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About

This is a second version of the pre-processing access layers script

see: https://github.com/BiologicalRecordsCentre/DECIDE-app/issues/141 for discussion



```{r load_pacakges}
library(osmextract)
library(sf)
library(rgdal)
library(raster)
library(dplyr)

library(leaflet)
```

```{r load_grids}
#UK 10km grids
uk_grid <- st_read('/data/data/DECIDE_constraintlayers/raw_data/UK_grids/uk_grid_10km.shp')
st_crs(uk_grid) <- 27700

# the big raster
raster100 <- raster::stack('/data/data/DECIDE_constraintlayers/environmental_data/100mRastOneLayer.grd')
#plot(raster100)
#get the CRS for when we project back to raster from data ramew
raster_crs <- st_crs(raster100)
raster100_df <- as.data.frame(raster100, xy=T, centroids=TRUE)[,1:2]
```





### Function for getting offline files for a 10km grid number

```{r}
#base location
base_location <- '/data/data/DECIDE_constraintlayers/raw_data/'

#folders
file_locations <- c(
  'CRoW_Act_2000_-_Access_Layer_(England)-shp/gridded_data_10km/',
  'OS_greenspaces/OS Open Greenspace (ESRI Shape File) GB/data/gridded_greenspace_data_10km/',
  #'SSSIs/gridded_data_10km/',
  'greater-london-latest-free/london_gridded_data_10km/',
  'rowmaps_footpathbridleway/rowmaps_footpathbridleway/gridded_data_10km/',
  #'RSPB_Reserve_Boundaries/gridded_data_10km/',
  'national_trust/gridded_data_10km/'
)

# function to get the data
get_offline_gridded_data <- function(grid_number){
  returned_files <- list()
  
  for (ii in 1:length(file_locations)){
    file_location <- file_locations[ii]
    all_grids <- list.files(paste0(base_location,file_location))
    
    right_file <- all_grids[grep(paste0("_",grid_number,".shp"), all_grids)]
    
    if (length(right_file)>0){
      #build file path
      file_path <- paste0(base_location,file_location,right_file)
      
      #load in file
      file_shp <- st_read(file_path, quiet = TRUE) %>% st_transform(4326)
      
      #add it to the object that this function returns
      returned_files[[length(returned_files)+1]] <- file_shp
    }
  }
  #return the files
  returned_files
}

#test <- get_offline_gridded_data(1500)
```

### Load in the OSM data

Vector translate operations - run once

Essentially this decompresses from the `.psm.pbf` file that we downloaded from geofabrik and by specifying extra tags that we can then SQL query. 

```{r}
#Define the vectortranslate options (see https://docs.ropensci.org/osmextract/articles/osmextract.html)

if(F){
  # ACCESS LAYERS
  #access lines
  
  vectortranslate_opts_access = c(
    "-select", "osm_id, highway, designation, footway, sidewalk",
    "-where", "highway IN ('footway', 'path', 'residential','unclassified','tertiary','sidewalk') OR designation IN ('public_footpath','byway_open_to_all_traffic','restricted_byway','public_bridleway','access_land') OR footway = 'sidewalk' OR sidewalk IN ('both','left','right')"
  )
  
  #do the vector translation
  oe_vectortranslate("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                     force_vectortranslate =T,
                     extra_tags = c("designation","footway","sidewalk"),
                     vectortranslate_options = vectortranslate_opts_access
                     )
  
  file.rename("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.gpkg",
              "/data/data/DECIDE_constraintlayers/raw_data/OSM/access_layers.gpkg")
  
  #NO GO AND WARNING AREAS
  #bad features lines (highway, landuse, aeroway,railway)
  vectortranslate_opts_avoid = c(
    "-select", "osm_id, highway, railway",
    "-where", "highway IN ('motorway','trunk') OR railway = 'rail'"
  )
  
  oe_vectortranslate("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                     force_vectortranslate =T,
                     extra_tags = c("railway"),
                     vectortranslate_options = vectortranslate_opts_avoid
                     )
  
  file.rename("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.gpkg",
              "/data/data/DECIDE_constraintlayers/raw_data/OSM/avoid_layers_lines.gpkg")
  
  
  #bad features areas(highway, landuse, aeroway,railway)
  vectortranslate_opts_avoid = c(
    "-select", "osm_id, landuse, aeroway",
    "-where", "landuse IN ('quarry','landfill','industrial','military') OR aeroway = 'aerodrome'"
  )
  
  oe_vectortranslate("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                     force_vectortranslate =T,
                     extra_tags = c("landuse", "aeroway"),
                     vectortranslate_options = vectortranslate_opts_avoid,
                     layer="multipolygons"
                     )
  
  #rename the file
  file.rename("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.gpkg",
              "/data/data/DECIDE_constraintlayers/raw_data/OSM/avoid_layers_area.gpkg")
  
  
  
  
  
  
  
  
  
  # gdal_utils(
  #   util = "vectortranslate",
  #   source = "/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
  #   destination = "/data/data/DECIDE_constraintlayers/raw_data/OSM/water_layers.gpkg",
  #   options = c(
  #     "-f", "GPKG", # output file format for GDAL < 2.3
  #     "-overwrite"
  #   ),
  #   quiet = F
  # )
  
  
  # WATER AREAS
  vectortranslate_opts_water = c(
    "-select", "osm_id, natural",
    "-where", "natural = 'water'"
  )
  
  oe_vectortranslate("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                     force_vectortranslate =T,
                     vectortranslate_options = vectortranslate_opts_water,
                     layer = 'multipolygons'
                     )
  
  file.rename("/data/data/DECIDE_constraintlayers/raw_data/OSM/geofabrik_south-yorkshire-latest.gpkg",
              "/data/data/DECIDE_constraintlayers/raw_data/OSM/water_layers.gpkg")


}
```

```{r OSM_queries}

#access lines
q1 <- "
SELECT * FROM 'lines' WHERE 
designation = 'public_footpath'
OR designation = 'public_bridleway'
OR designation = 'byway_open_to_all_traffic'
OR designation = 'restricted_byway'
OR designation = 'public_bridleway'
OR highway = 'footway'
OR highway = 'path'
OR highway = 'residential'
OR highway = 'unclassified'
OR highway = 'tertiary'
OR footway = 'sidewalk'
OR sidewalk = 'both'
OR sidewalk = 'left'
OR sidewalk = 'right'
"

#access polygons
q2 <- "
SELECT * FROM 'lines' WHERE
designation = 'access_land'
"


# no go lines
q3 <- "
SELECT * FROM 'lines' WHERE
railway = 'rail' OR
highway = 'motorway' OR
highway = 'trunk'
"

# no go, warning polys 
q4 <- "
SELECT * FROM 'multipolygons' WHERE
landuse2 = 'military' OR
aeroway2 = 'aerodrome' OR
landuse2 = 'quarry' OR
landuse2 = 'landfill'
"

# water polys
q5 <- "
SELECT * FROM 'polygons' WHERE
natural = 'water'
"



```

Proof of concept downloading data

```{r get_osm_data}

this_10k_grid <- uk_grid[1313,]$geometry #get the 10kgrid
this_10k_gridWGS84 <- st_transform(this_10k_grid, 4326) %>% st_as_text()# convert to WGS84

#
access_lines <- st_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/access_layers.gpkg",
                        layer = "lines",
                        query = q2,
                        wkt_filter = this_10k_gridWGS84
                        )

plot(access_lines)



#lines to avoid (motorways, railways)
avoid_lines <- st_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/avoid_layers.gpkg", 
                        query = q3,
                        wkt_filter = this_10k_gridWGS84
                        )

#areas to avoid or warn (military, landfill, industrial)
avoid_poly <- st_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/avoid_layers_area.gpkg",
                        #wkt_filter = this_10k_gridWGS84#,
                        #query = q4
                      )

#water areas
water_poly <- st_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/water_layers.gpkg")




access_lines %>% filter(highway == "trunk")

access_lines$highway %>% unique()

plot(access_lines,max.plot = 12)




m <- leaflet() %>%
    addTiles() %>%
    addPolylines(data=avoid_poly$geometry)
m


areas <- osmextract::oe_get(
  place = "South Yorkshire", 
  layer = "multipolygons", 
  query = "SELECT * FROM multipolygons WHERE natural = 'water'", 
  quiet = FALSE,
  force_download = T,
  force_vectortranslate = T
)

areas




```







Function that takes

Grid
loaded in gpkg file
raster

Set up to be able to be run as jobs

```{r}
assess_accessibility <- function(grid_number,grids){
  
  #load 10k grid
  this_10k_grid <- grids[grid_number,]$geometry #get the 10kgrid
  this_10k_gridWGS84 <- st_transform(this_10k_grid, 4326) # convert to WGS84
  
  #get the centroids of the the 100m grids within the 10k grid
  
  
  
  
  # load OSM data
  
  
  
  # Load offline data
  get_offline_gridded_data(grid_number)
  
  
  
  
}



```






























Graveyard

```{r}

oe_vectortranslate("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                   
                   force_vectortranslate =T,
                   extra_tags = c("designation","footway","sidewalk")
                   )



# test <- oe_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.gpkg",
#                 extra_tags = c("designation","footway","sidewalk"),
#                 query = q1,
#                 #boundary = this_10k_gridWGS84,
#                 #boundary_type = "clipsrc"
#                 )


test2 <- oe_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.gpkg")

access_lines <- oe_read("/data/data/DECIDE_constraintlayers/raw_data/OSM/great-britain-latest.osm.pbf",
                extra_tags = c("designation","footway","sidewalk"),
                query = q1)

```


```{r}


# 1. Download the data and skip gpkg conversion
oe_get("Portugal", download_only = TRUE, skip_vectortranslate = TRUE)

# 2. Define the vectortranslate options
my_vectortranslate = c(
  # SQL-like query where we select only the features where highway in (primary, secondary, tertiary)
  "-select", "osm_id, highway",
  "-where", "highway IN ('primary', 'secondary', 'tertiary')"
)

# 3. Convert and read-in
system.time({
  portugal1 = oe_get("Portugal", vectortranslate_options = my_vectortranslate)
})



```



